Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-07-08 manuscript/manuscript.Rmd
adds paragraph on caveats re RF

\begin{enumerate}[label=\textbf{\arabic*}]
        \item Rapid environmental change has generated growing interest in forecasts of future population trajectories. Traditional population models built with detailed demographic observations from one study site can address the impacts of environmental change at particular locations, but are difficult to scale up to the landscape and regional scales relevant to management decisions.
        \item An alternative is to build models using population-level data which are much easier to collect than individual-level data over broad spatial scales.
However, it is unknown whether models built using population-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.
        \item Here, we test the consequences of aggregating individual responses when forecasting the population states and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA. We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting skill and forecast horizons with and without the inclusion of climate covariates. For both models we used Bayesian ridge regression to weight the influence of climate covariates for optimal prediction. 
        \item In the absence of climate effects, we found no significant difference between the forecasting skill of models based on individual-level data and models based on population-level data. Climate effects were weak, but increased forecasting skill for two species. Increases in skill with climate covariates were similar between model types.
        \item For our focal species at this particular location, and using our particular statistical models, percent cover models generated forecasts as skillful as those from a demographic model. We conclude that models based on aggregated individual-level data offer a practical alternative to data-intensive demographic models when forecasting is the goal.
    \end{enumerate}

@@

>>>>

\begin{enumerate}[label=\textbf{\arabic*}]
        \item Rapid environmental change has generated growing interest in forecasts of future population trajectories. Traditional population models built with detailed demographic observations from one study site can address the impacts of environmental change at particular locations, but are difficult to scale up to the landscape and regional scales relevant to management decisions.
        \item An alternative is to build models using population-level data which are much easier to collect than individual-level data over broad spatial scales.
However, it is unknown whether models built using population-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.
        \item Here, we test the consequences of aggregating individual responses when forecasting the population states and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA. We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting skill and forecast horizons with and without the inclusion of climate covariates. For both models we used Bayesian ridge regression to weight the influence of climate covariates for optimal prediction. 
        \item In the absence of climate effects, we found no significant difference between the forecasting skill of models based on individual-level data and models based on population-level data. Climate effects were weak, but increased forecasting skill for two species. Increases in skill with climate covariates were similar between model types.
        \item For our focal species at this particular location, and using our particular statistical models, percent cover models generated forecasts as skillful as those from a demographic model. We conclude that models based on aggregated individual-level data offer a practical alternative to data-intensive demographic models when forecasting is the goal. Percent cover data are widely collected for many plant species, and several long-term time series already exist. Our results should inspire plant population modelers to exploit such data to forecast the impacts of climate change.
    \end{enumerate}

@@

Our comparison of two model types, one based on individual-level data and one based on percent cover data, should not be considered an exhaustive test.
There are several reasons why the percent cover-based model generated equally skillfull forecasts for our focal species, and several reasons why percent cover models may fail in other cases.
First, we had no cases where a climate covariate had a strong negative effect on one vital rate and a strong positive effect on a different vital rate (Fig. 3).
As noted by @Freckleton2011, complex age or stage structure can compromise predictions from density-structured models that aggregate over life-histories.
Interactions between a species' response to climate and life-history stages would also impact model forecasts because important climate effects could be averaged-over.
Second, our particular recruitment model is rather coarse and may therefore already average over important climate-influenced processes such as seed germination and establishment.
If we had been able to more precisely model the recruitment process, we might have discovered stronger climate effects -- effects that could be hard to detect from percent cover data.
As advocated by @Freckleton2011, knowledge of a species' population ecology should guide the modeling approach.
